Engineering the Servo Web Browser Engine using Rust

### 2015-10-23 icse16-servo/browsers.tex
ICFP reviewer comments

The set of memory buffers or graphical surfaces, called \emph{layers}, are then
transformed and composited together to form a final image for
presentation. Layerization is used to optimize interactive transformations
like scrolling and certain animations.

>>>>

The set of memory buffers or graphical surfaces, called \emph{layers}, are then
transformed and composited together to form a final image for
presentation.
These layers are then used to optimize interactive transformations, such as scrolling
and certain animations, by only redrawing the buffers that have changed and
otherwise simply re-composing the surface that has already been rendered into.

### 2015-10-23 icse16-servo/intro.tex
ICFP reviewer comments

%!TEX root = paper.tex
\section{Introduction}
\label{sec:intro}
The heart of a modern web browser is the browser engine, which is the code responsible
for loading, processing, evaluating, and rendering web content.
There are three major browser engine families:
\begin{enumerate}
\item Trident/Spartan, the engine in Internet Explorer~\cite{IE}
\item Webkit\cite{WEBKIT}/Blink, the engine in Safari~\cite{SAFARI}, Chrome~\cite{CHROME}, and Opera~\cite{OPERA}
\item Gecko, the engine in Firefox~\cite{FIREFOX}
\end{enumerate}
All of these engines have at their core many millions of lines of \Cplusplus{} code.
While the use of \Cplusplus{} has enabled all of these browsers to achieve excellent sequential
performance on a single web page, on mobile devices with lower processor speed but many
more processors, these browsers do not provide the same level of interactivity that they
do on desktop processors~\cite{parallelizing-web-pages,ZOOMM}.
Further, in an informal inspection of the critical security bugs in Gecko, we determined that
roughly 50\% of the bugs are use after free, out of range access, or related to integer
overflow.
The other 50\% are split between errors in tracing values from the JavaScript heap in the
\Cplusplus{} code and errors related to dynamically compiled code.

>>>>

%!TEX root = paper.tex
\section{Introduction}
\label{sec:intro}
The heart of a modern web browser is the browser engine, which is the code responsible
for loading, processing, evaluating, and rendering web content.
There are three major browser engine families:
\begin{enumerate}
\item Trident/Spartan, the engine in Internet Explorer~\cite{IE}
\item Webkit\cite{WEBKIT}/Blink, the engine in Safari~\cite{SAFARI}, Chrome~\cite{CHROME}, and Opera~\cite{OPERA}
\item Gecko, the engine in Firefox~\cite{FIREFOX}
\end{enumerate}
All of these engines have at their core many millions of lines of \Cplusplus{} code.
The use of \Cplusplus{} has enabled all of these browsers to achieve excellent sequential
performance on a single web page, particularly on desktop computers.
But, on mobile devices with lower processor speed but many more processors, these browsers
do not provide the same level of interactivity~\cite{parallelizing-web-pages,ZOOMM}.
Further, in an informal inspection of the critical security bugs in Gecko, we determined that
roughly 50\% of the bugs are use after free, out of range access, or related to integer
overflow.
The other 50\% are split between errors in tracing values from the JavaScript heap in the
\Cplusplus{} code and errors related to dynamically compiled code.

### 2015-10-23 icse16-servo/paper.tex
ICFP reviewer comments

We would like to thank Robert Clipsham, Valentin Gosu, David Herman, Tim Kuehn,
Lindsey Kuper, Ms2ger, and Ben Striegel for their comments on drafts of this paper.

>>>>

We would like to thank Robert Clipsham, Valentin Gosu, David Herman, Tim Kuehn,
Lindsey Kuper, Ms2ger, Ben Striegel, and previous anonymous reviewers for their
comments on drafts of this paper.

### 2015-10-23 icse16-servo/servo.tex
ICFP reviewer comments

\section{Servo}
\label{sec:servo}
A crucial test of Servo is performance --- Servo must be at least as fast
as other browsers at similar tasks to succeed, even if it provides additional memory safety.
\tblref{servo-perf} shows a preliminary comparison of the performance of the layout stage (described
in \secref{sec:browsers}) of rendering several web sites in Mozilla Firefox's Gecko engine compared
to Servo, taken on a modern MacBook Pro.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r | r | r}
      Site & Gecko & Servo 1 thread & Servo 4 threads \\
      \hline
      Reddit & 250 & 100 & 55  \\
      CNN & 105 & 50 & 35 \\
    \end{tabular}%
  \end{center}%
  \caption{Performance of Servo against Mozilla's Gecko rendering engine on the layout portion of some common sites.
  Times are in milliseconds, where lower numbers are better.}
  \label{servo-perf}
\end{table}

@@

Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This strategy is used by virtually all \Cplusplus{} compilers to implement templates, so it is proven and well-known
within systems programming.
Rust also follows this approach, although it improves on the ergonomics of \Cplusplus{} templates
by embedding serialized generic function ASTs within ``compiled'' binaries.

@@

One area for future improvement is related to allocations that are not owned by Rust itself.
Today, we simply wrap raw C pointers in \lstinline[language=Rust]{unsafe} blocks when we need to use a
custom memory allocator or interoperate with the SpiderMonkey JavaScript engine from Gecko.
We have implemented wrapper types and compiler plugins that restrict incorrect uses of these foreign values,
but they are still a source of bugs and one of our largest areas of unsafe code.

>>>>

\section{Servo}
\label{sec:servo}
A crucial test of Servo is performance --- Servo must be at least as fast
as other browsers at similar tasks to succeed, even if it provides additional memory safety.
\tblref{servo-perf} shows a preliminary comparison of the performance of the layout stage (described
in \secref{sec:browsers}) of rendering several web sites in Mozilla Firefox's Gecko engine compared
to Servo, taken on a modern MacBook Pro.
While these measurements only reflect layout performance, they show that Servo is already faster than Gecko with even a single thread, and achieves speedups with multiple threads.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r | r | r}
      Site & Gecko & Servo 1 thread & Servo 4 threads \\
      \hline
      Reddit & 250 & 100 & 55  \\
      CNN & 105 & 50 & 35 \\
    \end{tabular}%
  \end{center}%
  \caption{Performance of Servo against Mozilla's Gecko rendering engine on the layout portion of some common sites.
  Times are in milliseconds, where lower numbers are better.}
  \label{servo-perf}
\end{table}

@@

Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This strategy is used by virtually all \Cplusplus{} compilers to implement templates, so it is proven and well-known
within systems programming.
Rust also follows this approach, although it improves on the ergonomics of \Cplusplus{} templates
by embedding serialized generic function ASTs within ``compiled'' binaries so that library consumers can instantiate that AST directly instead of re-parsing the \Cplusplus{} header files for the templates.

@@

One area for future improvement is related to allocations that are not owned by Rust itself.
Today, we simply wrap raw C pointers in \lstinline[language=Rust]{unsafe} blocks when we need to use a
custom memory allocator or interoperate with the SpiderMonkey JavaScript engine from Gecko.
We have implemented wrapper types and compiler plugins that restrict incorrect uses of these foreign values,
but they are still a source of bugs and one of our largest areas of unsafe code.\footnote{https://blog.mozilla.org/research/2014/08/26/javascript-servos-only-garbage-collector/}
